Prestige as a Driving Force in Cultural Transmission

### 2022-03-30 ms/PrestigeBias.tex
fixes

\documentclass[12pt]{extarticle}
\usepackage{geometry}
\geometry{
a4paper,
total={170mm,257mm},
left=20mm,
top=20mm,
headheight=12pt
}

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage[unicode=true,colorlinks=true,urlcolor=Blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}

% unbreakable dashes
\usepackage[shortcuts]{extdash}
% footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

% less space before sections 
% \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
\usepackage{titlesec}
\titlespacing*{\section}
{0pt}{2ex plus 1ex minus .2ex}{2ex plus .2ex}
\titlespacing*{\subsection}
{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
\titlespacing*{\paragraph}
{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}

%SetFonts
% newtxtext+newtxmath
\usepackage{newtxtext} %loads helv for ss, txtt for tt
\usepackage{amsmath}
\usepackage[bigdelims]{newtxmath}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%SetFonts

@@

% Supplementary
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
}

% NatBib
%\usepackage[super,comma]{natbib}
\usepackage[round,colon,authoryear]{natbib}
%\use-package[numbers,super,sort&compress]{natbib}
\renewcommand{\bibsection}{}
%\renewcommand{\bibfont}{\small}

% math stuff
\DeclareMathOperator*{\E}{{\rm I\kern-.3em E}}
\newcommand*{\tr}{^\intercal}
\let\vec\mathbf
\newcommand{\matrx}[1]{{\left[ \stackrel{}{#1}\right]}}
\newcommand{\diag}[1]{\mbox{diag}\matrx{#1}}
\newcommand{\goesto}{\rightarrow}
\newcommand{\dspfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2} }
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}
\renewcommand\qedsymbol{} % no square at end of proof
\newcommand{\cl}{\mathbf{L}}
\newcommand{\cj}{\mathbf{J}}
\newcommand{\ci}{I}

% title page stuff
\title{Prestige as a Driving Force in Cultural Transmission}

\renewcommand\Affilfont{\small}
\author[1]{Saar Egozi}
\author[2,3,*]{Yoav Ram}
\affil[1]{School of Computer Science, Reichman University, Herzliya 4610101, Israel}
\affil[2]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv 6997801, Israel}
\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv 6997801, Israel}
\affil[*]{Corresponding author: yoav@yoavram.com, ORCID 0000-0002-9653-4458}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Copying role-models can be an efficient method for acquiring knowledge. A common bias when choosing a role-model to copy is success bias: copying whoever appears more successful. This bias depends on the performance of the role-model alone, with no other factors. We propose an additional bias that may be prevalent in cultural transmission: influence bias, in which role-model choice is affected by the number of individuals that have already copied each potential role-model. We combine success and influence bias together to a ``prestige bias'' and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations. We find analytic approximations to our stochastic model, facilitating further mathematical analysis and reducing the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading advantageous cultural trait, in both a constant and changing environment, which resemble Kimura's classical formulas.
These approximations show that success bias effectively plays the part of natural selection, whereas influence bias effectively reduces the population size. % TODO reduce or increase?
It also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a better description of human cultural transmission, 
especially in the last years where social networks are very popular. % TODO to discuss this last sentence
Further work is needed to test if this model could predict various phenomena in human cultural evolution when extended with the effects of selection and innovation.
\end{abstract}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

\paragraph{Cultural transmission.}
In cultural transmission, individuals transmit cultural traits (i.e., behavior, beliefs, norms) to one another, typically via learning and demonstrating \citep{transmissionVectorsBook}.
Examples for cultural traits in humans are behavioral patterns, such as personalities and habits, transmitted via both verbally and by observations. % TODO reference
Although cultural transmission is most common in humans, it is also observed in other animals such as chimpanzees \citep{chimpsPrestige, chimpsCopy}. % TODO but also birds, Cetacea (Whitehead, PNAS 2017)
In elephants, \citet{elepahntsRepo} showed that once a matriarch is removed from the group, the group's survival instincts are inferior and that ``the possession of enhanced discriminatory abilities by the oldest individual [matriarch] in a group can influence the social knowledge of the group as a whole.''
By playing audio recordings of African elephants, they showed that groups with a matriarch recognize and react better to hostile or friendly calls than groups without a matriarch.
\citet{fliesPaper} showed that choice of oviposition site in fruit flies is culturally transmitted: inexperiences flies that spent some time with experienced flies chose the same type of oviposition site even without directly observing this behavior. How the information is transmitted is still an open question, but it has been suggested that flies may use olfactory cues like other animals, such as rodents and bees.

\paragraph{Direction of transmission.}
Similar to genetic transmission, culturally transmitted traits can be transmitted from parents to offspring, and their effects of can be physiological rather than behavioral.
For example, parents can "teach" their children to be strong or tall, within some biological limits, by instructing them to maintain a specific diet and engage in physical activity.
Contrary to genetic transmission, cultural transmission can be non-vertical, that is, traits may be transmitted via social learning from non-parental individuals, and even unrelated individuals such as teachers, leaders, media, or any stranger that interacts with the learning individual.
Thus, cultural transmission may combine vertical transmission, where parents transmit to their offspring; oblique transmission, where adults transmit traits to unrelated offspring; and horizontal transmission, where peers from the same age cohort transmit to one another. 
Vertical transmission is also possible in the opposite direction: parents may copy traits from their offspring (e.g playing video games)\citep{transmissionVectorsBook,transmissionVectors}.  % TODO did they suggest video games or did they have some other trait? not sure I want to mention video games...

@@

\paragraph{Success bias.}
\citet{strategiesPaper} have conducted a tournament between learning strategies. Each strategy defines when individuals observe and copy from others, and when they engage in individual learning, in which an individual learns a cultural trait on his own. The best strategies had a high frequency of social learning relative to individual learning, even when the transmission error was almost 50\%. It is important to note that all of the strategies included some frequency of individual learning. 
%The authors suggested that success-biased social transmission contribute more to the general success of the population than individual learning. % TODO how is the last sentence related to the results of the tournament? did the best strategies use success-bias? it is not clear

% TODO do we need the following paragraph? it doesn't say much, and the paragraph after it covers "different type of success evaluation", so what is this paragraph doing?
%\citet{complexityPaper} define different types of transmission biases based on success. They define several types of role-model choice methods, all assuming that the copier correctly identifies the successful role-models.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Models and Methods}

@@

\subsection*{Continuous trait}
We follow the model of \citet{evolutionBook}, assuming only oblique transmission of the indicator trait and omitting the indirect trait in order to reduce model complexity. 
We consider a population of $N$ individuals, described by a single indicator trait with a continuous value.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright–Fisher model, we assume non-overlapping generations such that the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$. We assume the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

@@

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the indicator trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
% based on \cref{eq:copier_choose} is % TODO you refer an equation that did not yet appear!
\begin{equation}\label{eq:binary-model}
P_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + X\alpha'\beta(\hat{A})} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X} % TODO should it be G_{i,A} rather than P_i,A ?
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model.
% TODO what is alpha' ?
Initially, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. The rest of the details follow the continuous trait model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations such processes, summarized here and explained in detail below:
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$).
\item the expected number of copiers of each role-model equals its relative biased indicator value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the success weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated indicator value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$). 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the indicator values $A_j$ and the weights $\alpha_j$, as the estimation error is uniform. % TODO correct? anything else need to be assumed?
\end{enumerate}

\subsection*{Generalized Binomial Distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

\begin{result}[GBD approximation]
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

\begin{proof}
We denote $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ as the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ our of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1),
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citet{GBD}, which completes the proof.
%$Q_j(k,N)$ is the probability that $k$ out of $N$ copiers choose role-model $j$ at the end of the process, which by our previous notation is $k=K_{N,j}$. % TODO do you need this sentence?
\end{proof}

\begin{corollary}
$\E[K_{N,j}] = N \cdot G_{1,j}$.
\end{corollary}

\begin{proof}
\citet[eq.~(2.3)]{GBD} show that 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation).
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.
\end{proof}

\begin{corollary}
$\E[K_{Nj}] = \alpha_j \cdot \beta(A_j+e) / \overline{\alpha \cdot \beta(A+e)}$, where the averaging in the denominator is over the role-models index, $j$.
\end{corollary}

\begin{proof}
The initial prestige of role-model $j$ based on \cref{eq:prestige} is
\begin{equation}\label{eq:initial_prob}
G_{1,j} = \frac{\alpha_j\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N} \alpha_m\cdot\beta(A_m+e)} \;.
\end{equation}
The denominator of \cref{eq:initial_prob} can also be formulated as:
\begin{equation}\label{eq:denominator}
 \sum\limits_{m=1}^{N}\alpha_m\beta(A_m+e) = N \cdot \overline{\alpha \cdot \beta(A+e)} \;,
\end{equation}
where $\overline{\alpha\beta(A+e)}$ is the mean value of $\alpha_m\cdot\beta(A_m+e)$ for all $m$.
Using \cref{eq:denominator} we get, % TODO something is missing here? how did you get E[KNj] to relate to the previous two equations?
\begin{equation}
\E[K_{Nj}] = \alpha_j \cdot \beta(A_j+e) \bigg/ \overline{\alpha \cdot \beta(A+e)} \;,
\end{equation}
completing the proof.
\end{proof}

@@

% TODO what about a numerical validation for GBD?

\subsection*{Dirichlet-Multinomial Distribution}

\paragraph{P\'{o}lya urn model.}
This stochastic process consists of $N$ draws from an urn with an initial amount of colored balls of $M$ colors. When a ball is drawn, it is then placed back in the urn together with an additional new ball of the same color.
Let $\vec{U_i} = \{u_{i,1},u_{i,2},...,u_{i,M}\}$  where $u_{i,j}$ is the number of balls of the $j$-th color in the urn after $i$ draws.
Let $S_{i,j}=1$ when drawing a $j$-colored ball on the $i$-th draw, and $0$ otherwise. The probability that $S_{i,j}=1$ given $\vec{U_{i-1}}$ is
\begin{equation}\label{eq:polya}
\begin{split}
P(S_{i,j} = 1 \mid \vec{U_{i-1}}) = 
\frac{u_{i-1,j}}{\sum\limits_{m=1}^{M} u_{i-1,m}} = 
\frac{o_j + w_{i-1,j}}{\sum\limits_{m=1}^{M} o_m + w_{i-1,m}} = 
\frac{o_j + w_{i-1,j}}{i-1 + \sum\limits_{m=1}^{M} o_m} \;,
\end{split}
\end{equation}
where $o_j$ is the initial number of balls of the color $j$ in the urn, and $w_{i,j}$ is the cumulative number of new balls that were added to the urn after $i$ draws of the color $j$.
\\

\begin{result}[P{\'{o}}lya urn model]\label{result:polya}
The role-model choice process, $\big\{\vec{K}_i\big\}_{i=1}^N$, is equivalent to a \textit{P\'{o}lya urn model} if both trait estimation error and success weight are uniform in the population, $e=e_j$ and $\alpha=\alpha_j$ for all $j=1,\ldots,N$.
\end{result}

>>>>

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}          
\usepackage{graphicx}
\usepackage{amsthm, amsmath, amssymb}
\usepackage[font=small]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[displaymath, mathlines]{lineno}
\usepackage{natbib}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}

@@

% math stuff
\DeclareMathOperator*{\E}{{\rm I\kern-.3em E}}
\let\vec\mathbf
\newenvironment{conditions}
  {\par\vspace{\abovedisplayskip}\noindent\begin{tabular}{>{$}l<{$} @{${}={}$} l}}
  {\end{tabular}\par\vspace{\belowdisplayskip}}

% title page stuff
\author{
	Saar Egozi
	\and
	Yoav Ram
}

\begin{document}

\begin{titlepage}
\centering

\vspace{4cm}

{\LARGE\bfseries Prestige as a Driving Force in Cultural Transmission}

\vspace{2cm}

{\large\bfseries Saar Egozi}

\vspace{1cm}

{\large supervised by Dr. Yoav Ram}

\vspace{2cm}

\vfill

{\large Efi Arazi School of Computer Science, Reichman University}

\vspace{1cm}

{\itshape \today}

{\large saartk@gmail.com}
\end{titlepage}

\renewcommand\linenumberfont{\normalfont\small\sffamily}
\linenumbers
\modulolinenumbers[2]
\setcounter{secnumdepth}{0}
\tableofcontents

\clearpage

\section{Abstract}
Copying role-models can be an efficient method of acquiring knowledge. Here, we study the effects of prestige as a cultural transmission bias in cultural-evolution model. A common bias when choosing a role-model to copy is \textit{success bias}, i.e copying whoever appears more successful. This bias depends on the performance of the role-model alone, without other factors. We propose an additional bias that may be prevalent in cultural transmission: \textit{Influence bias}, which is evaluated by the number of individuals that have already copied a specific role-model. We combine these biases together to a \textit{prestige bias} and analyze its effects on the evolutionary dynamics of the population using mathematical analysis and stochastic simulations. We find analytic approximations to our model, facilitating further mathematical analysis and reducing the computational complexity of the simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
For a dichotomous cultural trait, we find  an approximation to the fixation probability and the fixation time of an invading advantageous cultural trait, in both a constant and changing environment, which resembles Kimura's classical formulas.
These approximations show that influence effectively reduces the population size. % TODO reduce or increase?
We also find that influence accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a better description of how human cultural transmission, 
especially in the last years where social networks are very popular. % TODO to discuss this last sentence
Further work is needed to test if this model could predict various phenomena in human cultural evolution when extended with the effects of selection and innovation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\paragraph{Cultural transmission.}
In cultural transmission, individuals transmit cultural traits (i.e., behavior, beliefs, norms) to one another, typically via learning and demonstrating \citep{transmissionVectorsBook}.
Examples for cultural traits in humans are behavioral patterns, such as personalities and habits, transmitted both verbally and by observations \citep{cultural_traits}.
Although cultural transmission is most common in humans, it is also observed in other animals such as chimpanzees \citep{chimpsPrestige, chimpsCopy}. % TODO but also birds, Cetacea (Whitehead, PNAS 2017)
In elephants, \citet{elepahntsRepo} showed that once a matriarch is removed from the group, the group's survival instincts are inferior and that ``the possession of enhanced discriminatory abilities by the oldest individual [matriarch] in a group can influence the social knowledge of the group as a whole.''
By playing audio recordings of African elephants, they showed that groups with a matriarch recognize and react better to hostile or friendly calls than groups without a matriarch.
\citet{fliesPaper} showed that choice of oviposition site in fruit flies is culturally transmitted: inexperiences flies that spent some time with experienced flies chose the same type of oviposition site even without directly observing this behavior. How the information is transmitted is still an open question, but it has been suggested that flies may use olfactory cues like other animals, such as rodents and bees.

\paragraph{Direction of transmission.}
Similar to genetic transmission, culturally transmitted traits can be transmitted from parents to offspring, and their effects of can be physiological rather than behavioral.
For example, parents can "teach" their children to be strong or tall, within some biological limits, by instructing them to maintain a specific diet and engage in physical activity.
Contrary to genetic transmission, cultural transmission can be non-vertical, that is, traits may be transmitted via social learning from non-parental individuals, and even unrelated individuals such as teachers, leaders, media, or any stranger that interacts with the learning individual.
Thus, cultural transmission may combine vertical transmission, where parents transmit to their offspring; oblique transmission, where adults transmit traits to unrelated offspring; and horizontal transmission, where peers from the same age cohort transmit to one another. 
Vertical transmission is also possible in the opposite direction: parents may copy traits from their offspring. \citep{transmissionVectorsBook,transmissionVectors}

@@

\paragraph{Success bias.}
Believed to be a main bias in nature for choosing a role-model for copying \citep{evolution_of_cultural_evolution}
\citet{strategiesPaper} have conducted a tournament between learning strategies based on that assumption.In their tournament, each strategy defines when individuals observe and copy from others, and when they engage in individual learning, in which an individual learns a cultural trait on his own. The best strategies had a high frequency of social learning relative to individual learning, even when the transmission error was almost 50\%. It is important to note that all of the strategies included some frequency of individual learning. 
%The authors suggested that success-biased social transmission contribute more to the general success of the population than individual learning. % TODO how is the last sentence related to the results of the tournament? did the best strategies use success-bias? it is not clear

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Models and Methods}

@@

\subsection{Continuous trait}
We follow the model of \citet{evolutionBook}, assuming only oblique transmission of the indicator trait and omitting the indirect trait in order to reduce model complexity. 
We consider a population of $N$ individuals, described by a single indicator trait with a continuous value.
Each generation, $N$ naive individuals, or copiers, choose an individual from the previous generation, or role-models, from which they will copy their trait. Similar to a Wright–Fisher model, we assume non-overlapping generations such that the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$. We assume the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

@@

\subsection{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the indicator trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
% based on \cref{eq:copier_choose} is % TODO you refer an equation that did not yet appear!
\begin{equation}\label{eq:binary-model}
P_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + X\alpha'\beta(\hat{A})} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X} % TODO should it be G_{i,A} rather than P_i,A ?
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model.
% TODO what is alpha' ?
Initially, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. The rest of the details follow the continuous trait model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\subsection{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations such processes, summarized here and explained in detail below:
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$).
\item the expected number of copiers of each role-model equals its relative biased indicator value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if $\alpha$ is uniform for all role-models ($\alpha=\alpha_k$ for $k=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated indicator value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$). 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the indicator values $A_j$ and the weights $\alpha_j$, as the estimation error is uniform. % TODO correct? anything else need to be assumed?
\end{enumerate}

\subsection{Generalized Binomial Distribution Approximation}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).

\paragraph{Proposition.} The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$.

\paragraph{Proof.} We denote $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ as the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ our of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1),
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citet{GBD}, which completes the proof.
%$Q_j(k,N)$ is the probability that $k$ out of $N$ copiers choose role-model $j$ at the end of the process, which by our previous notation is $k=K_{N,j}$. % TODO do you need this sentence?

\paragraph{Corollary 1.} $\E[K_{N,j}] = N \cdot G_{1,j}$.
\paragraph{Proof.} \citet[eq.~(2.3)]{GBD} show that 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation).
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.

\paragraph{Corollary 2: } $\E[K_{Nj}] = \alpha_j \cdot \beta(A_j+e) / \overline{\alpha \cdot \beta(A+e)}$, where the averaging in the denominator is over the role-models index, $j$.
\paragraph{Proof.} The initial prestige of role-model $j$ based on \cref{eq:prestige} is
\begin{equation}\label{eq:initial_prob}
G_{1,j} = \frac{\alpha_j\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N} \alpha_m\cdot\beta(A_m+e)} \;.
\end{equation}
The denominator of \cref{eq:initial_prob} can also be formulated as:
\begin{equation}\label{eq:denominator}
 \sum\limits_{m=1}^{N}\alpha_m\beta(A_m+e) = N \cdot \overline{\alpha \cdot \beta(A+e)} \;,
\end{equation}
where $\overline{\alpha\beta(A+e)}$ is the mean value of $\alpha_m\cdot\beta(A_m+e)$ for all $m$.
Using \cref{eq:denominator} we get, % TODO something is missing here? how did you get E[KNj] to relate to the previous two equations?
\begin{equation}
\E[K_{Nj}] = \alpha_j \cdot \beta(A_j+e) \bigg/ \overline{\alpha \cdot \beta(A+e)} \;,
\end{equation}
completing the proof.

@@

\subsection{Dirichlet-Multinomial Distribution Approximation}
\paragraph{Reminder: \textit{Pólya urn model}}  is a stochastic process that is defined as such: 
The process consists of $N$ draws from an urn with an initial amount of colored balls of $M$ colors. When a ball is drawn, it is then placed back in the urn together with an additional new ball of the same colour.\\
Let $\vec{U_i} = \{u_{i,1},u_{i,2},...,u_{i,M}\}$  where $u_{i,j}$ is the number of balls of the $j$-th color in the urn after $i$ draws.
Let $S_{i,j}=1$ when drawing a $j$ colored ball on the $i$-th draw, and $0$ otherwise. The probability that $S_{i,j}=1$ given $\vec{U_{i-1}}$ is:
\begin{equation}\label{polya}
\begin{split}
P(S_{i,j} = 1 | \vec{U_{i-1}}) & = \frac{u_{i-1,j}}{\sum\limits_{m=1}^{M} u_{i-1,m}} = \frac{o_j + w_{i-1,j}}{\sum\limits_{m=1}^{M} o_m + w_{i-1,m}}\\
 & = \frac{o_j + w_{i-1,j}}{i-1 + \sum\limits_{m=1}^{M} o_m}
\end{split}
\end{equation}
where $o_j$ is the initial number of balls of the colour $j$ in the urn, and $w_{i,j}$ is the number of new balls that were added to the urn after $i$ draws of the color $j$.

\paragraph{Proposition:} process $\big\{\vec{K}_i\big\}_{i=1}^N$ is equivalent to a \textit{Pólya urn model} when $e=e_i=e_j$ and $\alpha=\alpha_j=\alpha_i$ for all $i,j \in N$.
\paragraph{Proof:} 
We denote $\alpha'$ as the odds ratio between the weights of the indicator and the influence ($\alpha'=\frac{\alpha}{1-\alpha}$). 
Using \cref{eq:prestige} we get:
\begin{equation}\label{eq:copier_choose}
\begin{split}
G_{i,j} & = \frac{\alpha\cdot \beta(A'_j) + (1-\alpha) \cdot K_{i-1,j}}{W_i} \cdot \frac{1-\alpha}{1-\alpha} \\
&= \frac{\alpha'\beta(A'_j) + K_{i-1,j}}{\sum\limits_{m=1}^{N} \alpha'\beta(A'_m) + K_{i-1,m}}\\
& =\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{i-1 + \sum\limits_{m=1}^{N} \alpha'\beta(A'_m)}
\end{split}
\end{equation}
We see that \cref{polya} and \cref{eq:copier_choose} are equivalent when setting $M=N$, $o_j = \alpha'\beta(A'_j)$, $w_{i,j} = K_{i,j}$, completing the proof.
